\Large \bf Cyclone: High Availability for Persistent Key Value Stores

### 2017-08-10 paper/paper.tex
Tweak.
Updates.

These two performance bottlenecks illustrated in Figure~\ref{fig:problem} are
\emph{not independent}, they cannot be addressed in isolation. In particular,
durability semantics mean that the key-value store cannot respond to the client
until the operation has been persisted to the log, since it must be completed
even on a crash and subsequent recovery. Therefore, addressing the network
bottleneck is pointless without also addressing the storage bottleneck. Cyclone
shows how a small amount of non-volatile memory can be used to address both
bottlenecks.

@@

\section{Network}
\label{sec:network}
Cyclone uses the RAFT~\cite{raft} consensus protocol to replicate the log of
operations across machines. Cyclone provides guarantee to the client that any
operation responded
to is durable and its results \emph{available} as long as a majority of replicas
continue to function. Cyclone's failure model is \emph{fail-recovery},
where nodes that can recover from failure simply continue to participate in
the protocol. This
requires RAFT to persist every log entry \emph{before} it sends it out for
replication. Storage overheads are therefore not independent of network
overheads, and we must deal with both storage and networking overheads
simultaneously to improve the performance of replication.

@@

Software packet switches make a simple (usually stateless or
sometimes flow associated) decision for each packet. In contrast, running a
consensus protocol requires a relatively complex decision process for each
packet. The direction of flow of log entries depends on the role of the
replicas. Further, the assigned role to a replica can change on reconfiguration
after a failure. A replica must also reject packets from old configurations. In
Section~\ref{sec:dm} we cover the techniques we use in Cyclone to ameliorate
these overheads.

@@

RAFT requires that the log entry be persisted \emph{before} it is multicast out
to follower replicas. DPDK userspace NIC drivers operate in DDIO
mode~\cite{ddio} where the packet is directly written into the CPU cache rather
than first being DMAed into DRAM and then fetched by the CPU on demand. Since
the NVM is directly attached we need to persist the packet to it by executing a
cacheline flush ({\tt clflush}) instruction for every cacheline in the packet
and the pointer in the pointer buffer to persist these via the memory bus. This
is not too onerous a burden because we can use the newly introduced {\tt
  clflush-opt}~\cite{clflush_opt} instruction specifically intended to
efficiently flush to persistent memory without the overhead of the serialization
normally introduced by {\tt clflush}. This allows us to hit full memory
bandwidth on current generation platforms, a quantity in excess of 200 Gb/s per
core which is well above the near term speeds of commodity NICs. We execute a
single serializing {\tt sfence} before the sequence of {\tt clflushes} to make
sure any dirty cachelines due to header manipulation related writes from the CPU
are sent to cache.

Although RAFT is an efficient consensus protocol in the common case, the
protocol state machine still adds significant overhead to each packet, relative
to the time for the packet to flow in from the NIC and back out to the replicas.
We address this problem for the \emph{loaded} case using batching - treating a
whole sequence of client commands as a single RAFT log entry, while avoiding any
copies to group these packets together. Figure~\ref{fig:batching} illustrates
how this is done. We use a burst receive call available in the DPDK userspace
driver to receive a burst of client packets at a time. We then chain these
packets together and treat them as a single log entry from the perspective of
RAFT, amortizing the control plane overheads over the packets (at most 32 at a
time due to current driver limitations). Crucially, batching in Cyclone does not
involve a latency-throughput tradeoff like in many other
systems~\cite{ix-dataplane}. The batch receive call we use in DPDK returns
immediately with whatever number of packets is available, including zero. We
always flush the transmit buffer after every call to DPDK to transmit packets to
replicas. Therefore, we never tradeoff latency for throughput when batching.

@@

The scalable commutativity rule~\cite{scalable_commutativity} states that for a
shared memory program to scale in performance with concurrency it must also have
an interface that is commutative for common case invocations. We have also used
interface level commutativity to extract parallelism for the purpose of scaling
replication performance. Others have used the same commutativity to create
per-cpu data structures to scale KV store performance~\cite{flodb}.

>>>>

These two performance bottlenecks illustrated in Figure~\ref{fig:problem} are
\emph{not independent}, they cannot be addressed in isolation. In particular,
durability semantics mean that the key-value store cannot respond to the client
until the operation has been persisted to the log at a majority of replicas,
since it must be completed even on a crash and subsequent recovery. Therefore,
addressing the network bottleneck is pointless without also addressing the
storage bottleneck. Cyclone shows how a small amount of non-volatile memory can
be used to address both bottlenecks.

@@

\section{Network}
\label{sec:network}
Cyclone uses the RAFT~\cite{raft} consensus protocol to replicate the log of
operations across machines. Cyclone provides guarantee to the client that any
operation responded to is durable and its results \emph{available} as long as a
majority of replicas continue to function. Cyclone's failure model is
\emph{fail-recovery}, where nodes that can recover from failure simply continue
to participate in the protocol. This requires RAFT to persist every log entry
\emph{before} it sends it out for replication and before a follower replica
responds to the leader. Storage overheads are therefore not independent of
network overheads, and we must deal with both storage and networking overheads
simultaneously to improve the performance of replication.

@@

However, the flow of packets through a software packet switch is usually very
simple: the packet enters through an ethernet port and after a simple (usually
stateless) decision is sent out through a set of chosen ports.  In contrast, log
replication requires a complex protocol state machine to decide what to do with
the packet. In addition, unlike software switches we cannot simply forget a
packet after transmitting it. We must append it to the log and act on it after
consensus is reached by a majority quorum. In Section~\ref{sec:dm} we cover the
techniques we use in Cyclone to ameliorate these overheads.

@@

RAFT requires that the log entry be persisted \emph{before} it is multicast out
to follower replicas. DPDK userspace NIC drivers operate in DDIO
mode~\cite{ddio} where the packet is directly written into the CPU cache rather
than first being DMAed into DRAM and then fetched by the CPU on demand. We need
to persist the packet to directly attached NVM by executing a cacheline flush
({\tt clflush}) instruction for every cacheline in the packet and the pointer in
the pointer buffer to persist these via the memory bus. This is not too onerous
a burden because we can use the newly introduced {\tt
  clflush-opt}~\cite{clflush_opt} instruction specifically intended to
efficiently flush to persistent memory without the overhead of the serialization
normally introduced by {\tt clflush}. This allows us to hit full memory
bandwidth on current generation platforms, a quantity in excess of 200 Gb/s per
core which is well above the near term speeds of commodity NICs. We execute a
single serializing {\tt sfence} before the sequence of {\tt clflushes} to make
sure any dirty cachelines due to header manipulation related writes from the CPU
are sent to cache.

Although RAFT is an efficient consensus protocol in the common case, the
protocol state machine still adds significant overhead to each packet, relative
to the time for the packet to flow in from the NIC and back out to the replicas.
We address this problem for the \emph{loaded} case using batching - treating a
whole sequence of client commands as a single RAFT log entry, while avoiding any
copies to group these packets together. Figure~\ref{fig:batching} illustrates
how this is done. We use a burst receive call available in the DPDK userspace
driver to receive a burst of client packets at a time. We then chain these
packets together and treat them as a single log entry from the perspective of
RAFT, amortizing the control plane overheads over the packets (at most 32 at a
time due to current driver limitations). Crucially, batching in Cyclone does not
involve a latency-throughput tradeoff like in many other
systems~\cite{ix-dataplane}. The batch receive call we use in DPDK returns
immediately with whatever number of packets is available, including zero. We
always flush the transmit buffer after every call to DPDK to transmit packets to
replicas. Therefore, we never tradeoff latency for throughput when batching. We
also receive log entries in batches at follower replicas and return a single
acknowledgment for the entire batch, speeding up the progress of the protocol.

@@

%The scalable commutativity rule~\cite{scalable_commutativity} states that for a
%shared memory program to scale in performance with concurrency it must also have
%an interface that is commutative for common case invocations. We have also used
%interface level commutativity to extract parallelism for the purpose of scaling
%replication performance. Others have used the same commutativity to create
%per-cpu data structures to scale KV store performance~\cite{flodb}.
